---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2023-08-09"
---


# Estimation of population parameters in SRS

1. The following data gives the geographical area (in acres) under paddy for 58 villages. Draw sample of eight villages by using SRSWOR. Based on sample drawn find the estimate of 'average area per village under paddy' and estimate of its variance and 95\% confidence interval for 'average area per village under paddy'.

98, 19, 47, 137, 155, 270, 64, 69, 127, 292, 79, 81, 144, 104, 240, 273, 141, 56, 117, 201, 130, 58, 102, 170, 261, 158, 29, 102, 210, 189, 116, 46, 187, 101, 194, 93, 161, 222, 41, 127, 179, 223, 33, 114, 76, 96, 78, 88, 137, 114, 56, 108, 179, 318, 58, 58, 76, 272

$$\overline{X}\pm z_{\alpha/2}\times s\sqrt{\frac{1}{n}-\frac{1}{N}}$$

$$\sigma^2=\frac{1}{n}\sum{(x_i-\overline{x})}^2$$

$$S^2=\frac{1}{n-1}\sum{(x_i-\overline{x})}^2$$
$$\widehat{\sigma}^2=S^2=\frac{n}{n-1}var$$

```{r}
pop=c(98, 19, 47, 137, 155, 270, 64, 69, 127, 292, 79, 81, 144, 104, 240, 273, 141, 56, 117, 201, 130, 58, 102, 170, 261, 158, 29, 102, 210, 189, 116, 46, 187, 101, 194, 93, 161, 222, 41, 127, 179, 223, 33, 114, 76, 96, 78, 88, 137, 114, 56, 108, 179, 318, 58, 58, 76, 272)
N=length(pop)
n=8
alpha=0.05
X1=sample(pop,n,replace=F)
Xbar=mean(X1) # Sample Mean
S2x=(n/(n-1))*var(X1) # Sample Variance
z_alphaby2=qnorm(1-alpha/2)
Lower=Xbar-z_alphaby2*sqrt(S2x*(1/n-1/N))
Upper=Xbar+z_alphaby2*sqrt(S2x*(1/n-1/N))
cat("Confidenace Interval for pop mean = (",c(Lower,Upper),")")
cat("\nPopulation Mean",mean(pop), "Estimated Mean:",Xbar)
```

2. The following are the monthly expenditures of 25 households selected randomly by using SRSWOR from a village having N = 400 households.

1698, 1889, 1910, 1768, 1852, 1837, 1811, 1762, 1722, 1851, 1717, 1715, 1791, 1908, 1801, 1800, 1893, 1771, 1709, 1772, 1667, 1690, 1811, 1816, 1731.

Obtain an unbiased estimate of monthly average expenditure per household in the village. Also, provide an estimate of 98\% C.I. for the monthly average expenditure per household.

```{r}
n2=25
N2=400
X2=c(1698, 1889, 1910, 1768, 1852, 1837, 1811, 1762, 1722, 1851, 1717, 1715, 1791, 1908, 1801, 1800, 1893, 1771, 1709, 1772, 1667, 1690, 1811, 1816, 1731)
alpha2=0.02
X2bar=mean(X2) # Sample Mean
S2x2=(n2/(n2-1))*var(X2) # Sample Variance
z2_alphaby2=qnorm(1-alpha2/2)
Lower2=X2bar-z2_alphaby2*sqrt(S2x2*(1/n2-1/N2))
Upper2=X2bar+z2_alphaby2*sqrt(S2x2*(1/n2-1/N2))
sprintf("Estimated Mean: %1.3f",X2bar)
sprintf("Confidenace Interval for pop mean = (%1.3f,%1.3f)",Lower2,Upper2)
```

3. A SRS of 30 households was drawn from a city area containing 14,848 households. The number of persons per household in sample was as follows
5, 6, 3, 3, 2, 3, 3, 3, 4, 4, 3, 2, 7, 4, 3, 5, 4, 4, 3, 3, 4, 3, 3, 1, 2, 4, 3, 4, 2, 4

Estimate the total number of people in the area and compute the probability that this estimate is within $\pm$ 10 \% of the true value.

$$P(|X-\mu|\geq k\sigma)\leq \frac{1}{k^2} $$

$$N\overline{X}\sim Normal (N\mu,\frac{N^2\sigma^2}{n})$$
$$P(|N\overline{X}-N\mu|\geq k\frac{N\sigma}{\sqrt{n}})\leq \frac{1}{k^2} $$
$$P(|N\overline{X}-N\mu|\leq k\frac{N\sigma}{\sqrt{n}})\geq 1-\frac{1}{k^2} $$
$$k\frac{N\sigma}{\sqrt{n}}=10\% N\mu\rightarrow k=\frac{10\%\mu\sqrt{n}}{\sigma}$$

$$1-\frac{1}{k^2}=\frac{\sigma^2}{0.01\mu^2n}$$
```{r}
N3=14848
n3=30
X3=c(5, 6, 3, 3, 2, 3, 3, 3, 4, 4, 3, 2, 7, 4, 3, 5, 4, 4, 3, 3, 4, 3, 3, 1, 2, 4, 3, 4, 2, 4)
X3bar=mean(X3)
S2_3=(n3)/(n3-1)*var(X3)
Est_Tot_pop=N3*X3bar
sprintf("Estimate of Total Population: %1.3f",Est_Tot_pop)
sprintf("Probability that this estimate is within 10 percent of the true value:%1.4f",1-100*S2_3/(X3bar^2*n3))

```

4. The following table shows the weekly family income
$(X)$ and weekly family expenditure on food $(Y)$ in SRS of 30 low-income families. From this sample estimate the mean weekly expenditure on food per family and the percentage of income spent on food.

 Family No., 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15 
 Income (X), 620, 620, 870, 650, 580, 920, 880, 790, 830, 620, 630, 620, 600, 750, 900 
 Food Cost (Y), 143, 208, 227, 305, 412, 282, 242, 300, 242, 444, 134, 198, 294, 271, 222
 Family, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30 
 Income (X), 750, 690, 830, 850, 730, 660, 580, 770, 690, 650, 770, 690, 950, 770, 690 
 Food Cost (Y), 377, 226, 360, 206, 277, 259, 233, 398, 168, 378, 348, 287, 630, 195, 216 


```{r}
```